### 2019-01-22 manuscript.tex
A few edits
Added some new plots
made a few changes to the manuscript

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{9 Month Report}
\fancyfoot[LE,LO]{Page \thepage}

@@

\begin{titlepage}
	\centering
	%\includegraphics[width = number\textwidth]{image title}\par\vspace{1cm}}
	%this could be used to add the Aberdeen logo?
	{\scshape\Large The University of Aberdeen \par}
	\vspace{1cm}
	{\huge\bfseries Assymetry and the Absence of Patterns Do Not Encourage Optimal Decisions.\par}
	\vspace{1cm}
	{\Large\itshape Warren James\par}
	\vfill
	Supervised by\par
	Dr.~Amelia \textsc{Hunt}\par
	and\par
	Prof.~Ben \textsc{Tatler}\par
	\textbf{Word Count:} 12755
	\vfill
\end{titlepage}

@@

\paragraph{} The above sections have suggested ways in which the symmetry of the problem can be broken in order to cause one side to be more attractive than the other. However, one fundamental issue with this has already been mentioned; the pursuit of patterns \citep{wolford2004searching}. This predisposition to search for patterns is evident in the way in which it can interrupt the use of more sensible strategies \citep{wolford2004searching}, and also means we can "detect" patterns that simply do not exist \cite{yellott1969probability}. An experiment in the paper by \cite{yellott1969probability} asks people to make a simple guess as to whether one of two options will appear next. Unbeknownst to them, the option they pick will always be the correct answer. After having carried out this experiment, the participants were asked to explain how they made their choices and whether they had detected a pattern. Despite none existing, participants were able to produce some fairly elaborate patterns that also matched onto the choices they had made during the task.

@@

\paragraph{} Sequences of events such as the outcome of a coin toss or the outcome of a football match differ with respects to how outcome comes about. The coin toss relies on pure chance, whereas the series of football matches relies on the players' performance. This is what \cite{Ayton2004} argues causes one phenomenon to occur as opposed to the other. They argue that the \textit{Gambler's Fallacy} is more likely to occur when applied to mechanical events, and the \textit{Hot-Hand Illusion} will occur when there is an element of human performance. They tested this by presenting people with several sequence of outcomes that had a high rate of chance to change (i.e. the current observation will be different to the previous observation), or a low rate of change (i.e. high chance of there being repetitions). Participants were then to judge whether the sequence had come from a coin being flipped, or football team's performance in a tournament. Consistent with their hypothesis, their results indicated that people were more likely to attribute a sequence with a high rate of change (reflective of the \textit{Gambler's Fallacy}) to a more mechanical/random source.

\paragraph{} Interestingly, their results also appeared to show that when there was a $50\%$ chance of change (i.e. equal likelihood), participants were more likely to attribute this to human performance which suggests that no only are we prone to certain biases in certain contexts, but we also cannot recognise "true random" as being random \cite{Ayton2004}. This idea is not uncommon in Psychological literature and there are a wide range of studies that suggest humans struggle to produce random sequences \citep[see][for a review]{BARHILLEL1991428}. We also tend to believe that a random sequence should not contain long strings of repetitions and there should be a high rate of alternation between each potential event.

\paragraph{} Some have argued that people often appear to define randomness as being an absence of any discernable pattern \citep{lopes1987distinguishing}. For example, even though it's equally likely that flipping a coin 8 times will produce a sequence of exclusively heads as a sequence that has a mix of heads and tails (e.g. "HTTHHTHH"), people will still tend to believe that the latter example is "more random" than the other \citep{KAHNEMAN1972430}. In both these examples, the sequence is equally likely to occur, however, people tend to not recognise that a truly random sequence of coin flips could result in heads for every flip. \cite{GRIFFITHS201885} argues that this tendency to seek out a pattern results in us underestimating how stable a random sequence can be. In their paper \citep{GRIFFITHS201885}, they present a model that was tasked with deciding whether a sequence was randomly generated or generated by some sort of rule. They were attempting to match human performance in this task in that we tend to look for a higher rate of change before stating something is randomly produced \citep{Ayton2004}. In their model, they included run length and symmetry as factors in order to decide whether a sequence had been produced by a something random (e.g. a coin toss), or by other means. By incorporating this sensitivity to run length and symmetry, they were able to closely reproduce the average response of participants performing a similar task. This suggests that we have a great deal sensitivity to any perceptible patterns which in turn biases us towards believing that there is a pattern present even when there is none.

@@

\subsection*{Methods}
\addcontentsline{toc}{subsection}{Methods}
\subsubsection*{Participants}
\paragraph{} $10$\footnote{The aim was to have 18, but due to technical issues and an unusually high drop-out rate this was not reached. As such, any discussion of the results is preliminary at this time.} Participants (9 female) took part in this experiment, with an average age of 23 (SD = 3.2). Participants were recruited via word of mouth and posters. For participating in the experiment, participants were reimbursed \pounds10 for their time.

@@

\section*{Experiment 4: Unequal Reward}
\addcontentsline{toc}{section}{Experiment 3: Unequal Reward}

@@

\pararaph{} However, as discussed above in relation to the similarity of results obtained by both \cite{clarke2015failure}, who did not pay participants, and \cite{morvan2012human}, who did pay participants, both targets were not necessarilly equal in value for this experiment. In this task, participants were given 50p on each trial and were told they could choose to split the money equally across the potential targets (i.e. 25p/25p) or unequally (i.e. 40p/10p) prior to each trial. Using this setup, we are able to investigate whether participants that are more likely to opt for the unequal split would demonstrate a stronger tendency to stand closer to one hoop over the other.

>>>>

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{Breaking the Symmetry}
\fancyfoot[LE,LO]{Page \thepage}

@@

\begin{titlepage}
	\centering
	%\includegraphics[width = number\textwidth]{image title}\par\vspace{1cm}}
	%this could be used to add the Aberdeen logo?
	{\scshape\Large The University of Aberdeen \par}
	\vspace{1cm}
	{\huge\bfseries Asymetry and the Absence of Patterns Do Not Encourage Optimal Decisions.\par}
	\vspace{1cm}
	{\Large\itshape Warren James\par}
	\vfill
	Supervised by\par
	Dr.~Amelia \textsc{Hunt}\par
	and\par
	Prof.~Ben \textsc{Tatler}\par
	\textbf{Word Count:} 12755
	\vfill
\end{titlepage}

@@

\paragraph{} As mentioned above, one potential explanation as to why people opt for a stratey of probability matching may come from the search for patterns \citep{yellott1969probability}. This predisposition to search for patterns has been implicated as a reason why people fail to find more appropriate, and optimal solutions to  certain problems \citep{wolford2004searching}.  %The above sections have suggested ways in which the symmetry of the problem can be broken in order to cause one side to be more attractive than the other. However, one fundamental issue with this has already been mentioned; the pursuit of patterns \citep{wolford2004searching}. This predisposition to search for patterns is evident in the way in which it can interrupt the use of more sensible strategies \citep{wolford2004searching}, and also means we can "detect" patterns that simply do not exist \cite{yellott1969probability}. An experiment in the paper by \cite{yellott1969probability} asks people to make a simple guess as to whether one of two options will appear next. Unbeknownst to them, the option they pick will always be the correct answer. After having carried out this experiment, the participants were asked to explain how they made their choices and whether they had detected a pattern. Despite none existing, participants were able to produce some fairly elaborate patterns that also matched onto the choices they had made during the task.

@@

\paragraph{} Sequences of events such as the outcome of a coin toss or the outcome of a football match differ with respects to how each outcome comes about. The coin toss relies on pure chance, whereas the series of football matches relies on the players' performance. This is what \cite{Ayton2004} argues causes one phenomenon to occur as opposed to the other. They argue that the \textit{Gambler's Fallacy} is more likely to occur when applied to mechanical events, and the \textit{Hot-Hand Illusion} will occur when there is an element of human performance. They tested this by presenting people with several sequence of outcomes that had a high rate of chance to change (i.e. the current observation will be different to the previous observation), or a low rate of change (i.e. high chance of there being repetitions). Participants were then to judge whether the sequence had come from a coin being flipped, or football team's performance in a tournament. Consistent with their hypothesis, their results indicated that people were more likely to attribute a sequence with a high rate of change (reflective of the \textit{Gambler's Fallacy}) to a more mechanical/random source.

\paragraph{} Interestingly, their results also appeared to show that when there was a $50\%$ chance of change (i.e. equal likelihood), participants were more likely to attribute this to human performance which suggests that not only are we prone to certain biases in certain contexts, but we also cannot recognise "true random" as being random \citep{Ayton2004}. This idea is not uncommon in Psychological literature and there are a wide range of studies that suggest humans struggle to produce random sequences \citep[see][for a review]{BARHILLEL1991428}. We also tend to believe that a random sequence should not contain long strings of repetitions and there should be a high rate of alternation between each potential event.

\paragraph{} Some have argued that people often appear to define randomness as being an absence of any discernable pattern \citep{lopes1987distinguishing}. For example, even though it's equally likely that flipping a coin 8 times will produce a sequence of exclusively heads as a sequence that has a mix of heads and tails (e.g. "HTTHHTHH"), people will still tend to believe that the latter example is "more random" than the other \citep{KAHNEMAN1972430}. In both these examples, the sequence is equally likely to occur, however, people tend to not recognise that a truly random sequence of coin flips could result in heads for every flip. \cite{GRIFFITHS201885} argues that this tendency to seek out a pattern results in us underestimating how stable a random sequence can be. In their paper \citep{GRIFFITHS201885}, they present a model that was tasked with deciding whether a sequence was randomly generated or generated by some sort of rule. They were attempting to match human performance in this task in that we tend to look for a higher rate of change before stating something is randomly produced \citep{Ayton2004}. In their model, they included run length and symmetry as factors in order to decide whether a sequence had been produced by a something random (e.g. a coin toss), or by other means. By incorporating this sensitivity to run length and symmetry, they were able to closely reproduce the average response of people performing a similar task. This suggests that we have a great deal sensitivity to any perceptible patterns which in turn biases us towards believing that there is a pattern present even when there is none.

@@

\subsection*{Methods}
\addcontentsline{toc}{subsection}{Methods}
\subsubsection*{Participants}
\paragraph{} $16$ Participants (2 male) took part in this experiment, with an average age of 21 (SD = 9.7). Participants were recruited via word of mouth and posters. For participating in the experiment, participants were reimbursed \pounds10 for their time.

@@

\section*{Experiment 4: Unequal Reward}
\addcontentsline{toc}{section}{Experiment 4: Unequal Reward}

@@

\paragraph{} However, as discussed above in relation to the similarity of results obtained by both \cite{clarke2015failure}, who did not pay participants, and \cite{morvan2012human}, who did pay participants, both targets were not necessarilly equal in value for this experiment. In this task, participants were given 50p on each trial and were told they could choose to split the money equally across the potential targets (i.e. 25p/25p) or unequally (i.e. 40p/10p) prior to each trial. Using this setup, we are able to investigate whether participants that are more likely to opt for the unequal split would demonstrate a stronger tendency to stand closer to one hoop over the other.
